External Patch Group Prior Guided Internal Prior Learning for Real Image Denoising

### 2016-10-26 CVPR2017_Review_Unified.tex
add

Natural images often demonstrate repepitive local patterns, this nonlocal self-similarity (NSS) property is a key successful factor for many image denoising methods \cite{nlm,bm3d,lssc,ncsr,wnnm,pgpd}. In this section, we formulate the Patch Group prior learned on natural color images. Similar to \cite{pgpd}, the patch group (PG) is defined as a group of similar patches to the local patch. The patch group mean is destracted, and hence different groups patches can share similar PGs. In this way, the space natural image patches to be modeled is largely reduced. In this work, each local patch extracted from RGB images is of size $p\times p \times 3$. Then we search the $M$ most similar patches $\{\mathbf{x}_{m}\}_{m=1}^{M}$ around each local patch through Euclidean distance, in a local window of size $W\times W$. The $\mathbf{x}_{m}\in \mathbb{R}^{3p^{2}\times1}$ is a patch vector formed by combining the 3 patch vectors (of size $p^{2}\times 1$) in R, G, B channels. The mean vector of this PG is $\boldsymbol{\upmu}=\frac{1}{M}\sum_{m=1}^{M}\mathbf{x}_{m}$, and the group mean subtracted PG is defined as $\mathbf{\overline{X}}\triangleq \{\mathbf{\overline{x}}_{m}=\mathbf{x}_{m}-\boldsymbol{\upmu}\}, m=1,...,M$. Assume we have extracted $N$ PGs from a set of external natural images, and the $n$-th PG is defined as $\mathbf{\overline{X}}_{n}\triangleq \{\mathbf{\overline{x}}_{n,m}\}_{m=1}^{M}, n=1,...,N$. We employ the Gaussian Mixture Model (GMM) to learn the external patch group based NSS prior. In this model, the likelihood of the $n$-th PG $\{\mathbf{\overline{X}}_{n}\}$ can be calculated as
\vspace{-2mm}
\begin{equation}\label{equ1}\vspace{-2mm}
P(\mathbf{\overline{X}}_{n})  = \sum\nolimits_{k=1}^{K}\pi_{k}\prod\nolimits_{m=1}^{M}\mathcal{N}(\mathbf{\overline{x}}_{n,m}|\boldsymbol{\upmu}_{k},\mathbf{\Sigma}_{k}),
\end{equation}
where $K$ is the number of Gaussians and the parameters $\pi_{k}$, $\boldsymbol{\upmu}_{k}$, $\mathbf{\Sigma}_{k}$ are mixture weight, mean vector, and covariance matrix of the $k$-th Gaussian, respectively. By assuming that all the PGs are independently sampled, the overall objective log-likelihood function is
\vspace{-2mm}
\begin{equation}\label{equ2}\vspace{-2mm}
\begin{split}
\ln\mathcal{L}=\sum_{n=1}^{N} \ln(\sum_{k=1}^{K}\pi_{k}\prod_{m=1}^{M}\mathcal{N}(\mathbf{\overline{x}}_{n,m}|\boldsymbol{\upmu}_{k},\mathbf{\Sigma}_{k})).
\end{split}
\end{equation} 
We maximize the above objective function via EM algorithm \cite{em} and finally obtain the GMM model with learned parameters as the learned external PG prior. Similar to \cite{pgpd}, the mean vector of each cluster is natural zeros, i.e., $\boldsymbol{\upmu}_{k}=\mathbf{0}$.

@@

Given a real noisy image, we extract the noisy PGs and corresponding mean vectors. Each mean substracted PG is defined as $\mathbf{\overline{Y}}\triangleq \{\mathbf{\overline{x}}_{m}\}_{m=1}^{M}$. Noted that, different from the external PGs, the mean vectors of the internal noisy PGs are saved for recovering. For adaptivity, we project the PG $\mathbf{\overline{Y}}$ into its most suitable Gaussian component (subspace) of the GMM learned on external PGs. The subspace most suitable for $\mathbf{\overline{Y}}$ is selected by firstly calculating the posterior probability of "$\mathbf{\overline{Y}}$ belonging to the $k$th Gaussian component":
\vspace{-3mm}
\begin{equation}\label{equ3}\vspace{-2mm}
P(k|\mathbf{\overline{Y}})=\frac{\prod_{m=1}^{M}\mathcal{N}(\mathbf{\overline{y}}_{m}|\mathbf{0},\mathbf{\Sigma}_{k})}{\sum_{l=1}^{K}\prod_{m=1}^{M}\mathcal{N}(\mathbf{\overline{y}}_{m}|\mathbf{0},\mathbf{\Sigma}_{l})},
\end{equation}
and then choosing the component with the maximum A-posteriori (MAP) probability $\ln P(k|\mathbf{\overline{Y}})$.

@@

Now each internal noisy PG has been assigned to its most suitable Gaussian component $\mathcal{N}(\mathbf{0},\mathbf{\Sigma}_{k})$ (subspace). In order to recover the latent clean PG, we employ the sparse coding framework under which we need a suitable dictionary for coding. As we know, given a dictionary $\mathbf{D}$, its \emph{mutual incoherence} $\mu(\mathbf{D})$ \cite{donoho2003optimally}, which is difined as
\vspace{-2mm}
\begin{equation}\label{equ4}
\vspace{-2mm}
\mu(\mathbf{D})=\max_{i~=j}\frac{|\mathbf{d}_{i}^{T}\mathbf{d}_{j}|}{\|\mathbf{d}_{i}\|_{2}\|\mathbf{d}_{j}\|_{2}},
\end{equation}
is a measure of quality of dictionary. Since the \emph{mutual incoherence} of a orthogonal dictionary is $0$, its quality is often better than non-orthogonal dictionaries.

external subspace through covariance matrix via singular value decomposition (SVD) 
\begin{equation}\label{equ4}
\mathbf{\Sigma}=\mathbf{U}_{e}\mathbf{S}_{e}\mathbf{V}_{e}^{T}.
\end{equation}
can learn for each subspace an internal orthogonal dictionary adaptive to the noisy PGs. For notation simplicity, we ignore the index of subspace $k$.

\begin{equation}\label{equ5}
\begin{split}
&\min_{\mathbf{D}_{i}\in \mathbb{R}^{3p^2\times r},\textbf{A}\in\mathbb{R}^{3p^2\times MN}}
\|\mathbf{Y}-[\mathbf{D}_{e}\ \mathbf{D}_{i}]\mathbf{A}\|_{F}^{2}+\lambda\|\mathbf{A}\|_{1}
\\
&
\text{s.t.}
\quad
\mathbf{D}_{i}^{T}\mathbf{D}_{i} = \mathbf{I}_{r},\ \mathbf{D}_{e}^{T}\mathbf{D}_{i} = \mathbf{0},
\end{split}
\end{equation}
The singular vectors capture the statistical structures of NSS variations in natural images, while the singular values in $\mathbf{S}$ represent the significance of these singular vectors. Fig.\ 4 shows the singular vectors for one Gaussian component.

>>>>

Natural images often demonstrate repepitive local patterns, this nonlocal self-similarity (NSS) property is a key successful factor for many image denoising methods \cite{nlm,bm3d,lssc,ncsr,wnnm,pgpd}. In this section, we formulate the Patch Group prior learned on natural color images. Similar to \cite{pgpd}, the patch group (PG) is defined as a group of similar patches to the local patch. The patch group mean is destracted, and hence different groups patches can share similar PGs. In this way, the space natural image patches to be modeled is largely reduced.

In this work, each local patch extracted from RGB images is of size $p\times p \times 3$. Then we search the $M$ most similar patches $\{\mathbf{x}_{m}\}_{m=1}^{M}$ around each local patch through Euclidean distance, in a local window of size $W\times W$. The $\mathbf{x}_{m}\in \mathbb{R}^{3p^{2}\times1}$ is a patch vector formed by combining the 3 patch vectors (of size $p^{2}\times 1$) in R, G, B channels. The mean vector of this PG is $\boldsymbol{\upmu}=\frac{1}{M}\sum_{m=1}^{M}\mathbf{x}_{m}$, and the group mean subtracted PG is defined as $\mathbf{\overline{X}}\triangleq \{\mathbf{\overline{x}}_{m}=\mathbf{x}_{m}-\boldsymbol{\upmu}\}, m=1,...,M$. Assume we have extracted $N$ PGs from a set of external natural images, and the $n$-th PG is defined as $\mathbf{\overline{X}}_{n}\triangleq \{\mathbf{\overline{x}}_{n,m}\}_{m=1}^{M}, n=1,...,N$. We employ the Gaussian Mixture Model (GMM) to learn the external patch group based NSS prior. In this model, the likelihood of the $n$-th PG $\{\mathbf{\overline{X}}_{n}\}$ can be calculated as
\vspace{-2mm}
\begin{equation}\label{equ1}\vspace{-2mm}
P(\mathbf{\overline{X}}_{n})  = \sum\nolimits_{k=1}^{K}\pi_{k}\prod\nolimits_{m=1}^{M}\mathcal{N}(\mathbf{\overline{x}}_{n,m}|\boldsymbol{\upmu}_{k},\mathbf{\Sigma}_{k}),
\end{equation}
where $K$ is the number of Gaussians and the parameters $\pi_{k}$, $\boldsymbol{\upmu}_{k}$, $\mathbf{\Sigma}_{k}$ are mixture weight, mean vector, and covariance matrix of the $k$-th Gaussian, respectively. By assuming that all the PGs are independently sampled, the overall objective log-likelihood function is
\vspace{-2mm}
\begin{equation}\label{equ2}\vspace{-2mm}
\begin{split}
\ln\mathcal{L}=\sum_{n=1}^{N} \ln(\sum_{k=1}^{K}\pi_{k}\prod_{m=1}^{M}\mathcal{N}(\mathbf{\overline{x}}_{n,m}|\boldsymbol{\upmu}_{k},\mathbf{\Sigma}_{k})).
\end{split}
\end{equation} 
We maximize the above objective function via EM algorithm \cite{em} and finally obtain the GMM model with learned parameters. Similar to \cite{pgpd}, the mean vector of each cluster is natural zeros, i.e., $\boldsymbol{\upmu}_{k}=\mathbf{0}$.

Now, we have clustered the PGs extracted from external clean images into $K$ Gaussians or subspaces. For notation simplicity, we ignore the index of subspace $k$. To better characterize each subspace, we perform singular value decomposition (SVD) on the covariance matrix:
\vspace{-2mm}
\begin{equation}\label{equ3}\vspace{-2mm}
\mathbf{\Sigma}=\mathbf{U}_{e}\mathbf{S}_{e}\mathbf{V}_{e}^{T}.
\end{equation}
The singular vectors in $\mathbf{U}_{e}$ are employed as the external orthogonal dictionary to guide the internal orthogonal dictionary learning (the singular values are employed as prior weights for sparse coding which will be discussed in Section 4). For a dictionary $\mathbf{D}$, its \emph{mutual incoherence} $\mu(\mathbf{D})$ \cite{donoho2003optimally} difined by
\vspace{-2mm}
\begin{equation}\label{equ4}
\vspace{-2mm}
\mu(\mathbf{D})=\max_{i~=j}\frac{|\mathbf{d}_{i}^{T}\mathbf{d}_{j}|}{\|\mathbf{d}_{i}\|_{2}\|\mathbf{d}_{j}\|_{2}}
\end{equation}
is a measure of quality of dictionary (lower is better). The dictionary $\mathbf{U}_{e}$ always has $0$
\emph{mutual incoherence} and hence better quality than non-orthogonal dictionaries.

@@

Given a real noisy image, we extract the noisy PGs and corresponding mean vectors. Each mean substracted PG is defined as $\mathbf{\overline{Y}}\triangleq \{\mathbf{\overline{x}}_{m}\}_{m=1}^{M}$. Noted that, different from the external PGs, the mean vectors of the internal noisy PGs are saved for recovering. For adaptivity, we project the PG $\mathbf{\overline{Y}}$ into its most suitable Gaussian component (subspace) of the GMM learned on external PGs. The subspace most suitable for $\mathbf{\overline{Y}}$ is selected by firstly calculating the posterior probability of "$\mathbf{\overline{Y}}$ belonging to the $k$th Gaussian component":
\vspace{-2mm}
\begin{equation}\label{equ5}\vspace{-2mm}
P(k|\mathbf{\overline{Y}})=\frac{\prod_{m=1}^{M}\mathcal{N}(\mathbf{\overline{y}}_{m}|\mathbf{0},\mathbf{\Sigma}_{k})}{\sum_{l=1}^{K}\prod_{m=1}^{M}\mathcal{N}(\mathbf{\overline{y}}_{m}|\mathbf{0},\mathbf{\Sigma}_{l})},
\end{equation}
and then choosing the component with the maximum A-posteriori (MAP) probability $\ln P(k|\mathbf{\overline{Y}})$.

@@

Now all internal noisy PGs have been assigned to their corresponding most suitable Gaussians or subspaces in $\{\mathcal{N}(\mathbf{0},\mathbf{\Sigma}_{k})\}_{k=1}^{K}$. For each subspace, we employ its external orthogonal dictionary $\mathbf{D}_{e}$ to guide the internal orthogonal dictionary learning. The learning is performed under the sparse coding framework (along with which the denoising of real noisy image is simultaneously done) as follows:
\vspace{-2mm}
\begin{equation}\label{equ6}\vspace{-2mm}
\begin{split}
&\min_{\mathbf{D}_{i}\in \mathbb{R}^{3p^2\times r},\textbf{A}\in\mathbb{R}^{3p^2\times MN}}
\|\mathbf{Y}-[\mathbf{D}_{e}\ \mathbf{D}_{i}]\mathbf{A}\|_{F}^{2}+\lambda\|\mathbf{A}\|_{1}
\\
&
\text{s.t.}
\quad
\mathbf{D}_{i}^{T}\mathbf{D}_{i} = \mathbf{I}_{r},\ \mathbf{D}_{e}^{T}\mathbf{D}_{i} = \mathbf{0},
\end{split}
\end{equation}
The singular vectors capture the statistical structures of NSS variations in natural images, while the singular values in $\mathbf{S}$ represent the significance of these singular vectors. Fig.\ 4 shows the singular vectors for one Gaussian component.
